---
title: "Week9_2021_Feb23"
author: "Liam D. Bailey"
date: "23/02/2021"
output:
  html_document:
  theme: paper
highlight: kate
editor_options:
  chunk_output_type: console
---

Set options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.showtext = TRUE, fig.retina = 1)
options(scipen = 200)
```

Load data

```{r data}

tuesdata <- tidytuesdayR::tt_load(2021, week = 9)

employed <- tuesdata$employed
earn     <- tuesdata$earn

```

Load packages

```{r packages, message = FALSE}

library(dplyr)
library(ggplot2)
library(showtext)

```

Skim the earnings data

```{r}

skimr::skim(earn)

```

There are annoying categories for race, age, and sex. Sex includes 'both sexes', while race contains 'all races'. Age has many different, mutually exclusive categories (e.g. 16 years and over & 16 to 24). We can just use '16 years and over' because we are not using age.

```{r}

earn %>% 
  filter(year == 2010 & quarter == 1 & race == "Asian" & sex == "Men")

```

Consider black, asian and white employees (for now joining sexes together). Note, we are excluding Hispanic here, because it doesn't fit neatly within the 'race' categories.

```{r}

plot_data <- earn %>% 
  filter(sex == "Both Sexes" & race != "All Races" & age == "16 years and over") %>% 
  #Create a time values (combine year and quarter)
  mutate(time = year + (quarter-1)/4)

```

```{r}

ggplot(data = plot_data) +
  geom_area(aes(x = time, y = median_weekly_earn, fill = race), colour = "black") +
  scale_fill_discrete(name = "Race") +
  labs(x = "", y = "Median weekly income (USD$)") +
  theme_classic() +
  theme(legend.position = "bottom")

ggsave(filename = here::here("./plots/2021/Week9/start_plot.png"), height = 7, width = 9, dpi = 600, scale = 0.75)

```

Show in bars so it's clearer which year is which

```{r}

ggplot(data = plot_data) +
  geom_col(aes(x = time, y = median_weekly_earn, fill = race),
           colour = "black", alpha = 0.5) +
  geom_col(data = plot_data %>% 
             mutate(median_weekly_earn = dplyr::case_when(time == 2020.75 ~ median_weekly_earn,
                                                          TRUE ~ 0)),
           aes(x = time, y = median_weekly_earn, fill = race),
           colour = "black") +
  scale_fill_discrete(name = "Race") +
  labs(x = "", y = "Median weekly income (USD$)") +
  theme_classic() +
  theme(legend.position = "bottom")

ggsave(filename = here::here("./plots/2021/Week9/start_plot_bars.png"), height = 7, width = 9, dpi = 600, scale = 0.75)

```

This is fine, but what about the inequality between the different racial groups? We'll create a measure of inequality that is the ratio of median weekly income between the highest and lowest earning group.

```{r}

plot_data_inequality <- plot_data %>% 
  group_by(time) %>% 
  summarise(income_ratio = min(median_weekly_earn)/max(median_weekly_earn))

```

```{r}

ggplot(data = plot_data_inequality) +
  geom_line(aes(x = time, y = income_ratio), colour = "#193D58", size = 1) +
  scale_x_continuous(breaks = seq(2010, 2021, 1)) +
  labs(x = "", y = "Ratio between highest and lowest earning race") +
  theme_classic()

ggsave(filename = here::here("./plots/2021/Week9/inequality_allsex.png"), height = 7, width = 9, dpi = 600, scale = 0.75)

```

Interesting, look at this for males and females separately.

```{r}

inequality_by_sex <- earn %>% 
  filter(sex != "Both Sexes" & race != "All Races" & age == "16 years and over") %>% 
  #Create a time values (combine year and quarter)
  mutate(time = year + quarter/4) %>% 
  #Ratio between highest and lowest earning race at each time point for each sex
  group_by(sex, time) %>% 
  summarise(income_ratio = min(median_weekly_earn)/max(median_weekly_earn))

#Add sex symbols at last point
inequality_by_sex %>% 
  group_by(sex) %>% 
  slice(n()) %>% 
  ungroup() %>% 
  mutate(time = time + 0.5,
         sex_symbol = c("\u2642", "\u2640")) -> sex_symbol

ggplot() +
  geom_line(data = inequality_by_sex, aes(x = time, y = income_ratio, colour = sex), size = 0.75) +
  labs(x = "", y = "Ratio between \n highest and lowest earning race") +
  geom_hline(aes(yintercept = 1), lty = 2) +
  geom_text(aes(x = mean(inequality_by_sex$time), y = 0.99, label = "Income parity"), size = 5) +
  geom_text(data = sex_symbol, aes(x = time, y = income_ratio, label = sex_symbol), size = 10) +
  scale_colour_manual(name = "Sex", values = c("#266742", "#3255A2")) +
  scale_x_continuous(breaks = seq(2010, 2021, 1)) +
  scale_y_continuous(limits = c(NA, 1)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_text(vjust = 0.1),
        axis.title.y = element_text(vjust = 0.9))

```

Add a rolling mean to account for temporal stochasticity.

```{r}

rolling_mean <- inequality_by_sex %>% 
  group_by(sex) %>% 
  mutate(lag1 = lag(income_ratio),
         lead1 = lead(income_ratio)) %>% 
  rowwise() %>% 
  mutate(rollingmean = mean(c(income_ratio, lag1, lead1)))

ggplot() +
  geom_line(data = inequality_by_sex, aes(x = time, y = income_ratio, group = sex),
            colour = "grey50", size = 0.75, alpha = 0.5) +
  geom_line(data = rolling_mean, aes(x = time, y = rollingmean, colour = sex), size = 1) +
  labs(x = "", y = expression("Less equal" %<-% "         " %->% "More equal"),
       title = "Racial income inequality in the USA",
       subtitle = "Ratio between the income of the lowest and highest earning racial group",
       caption = "Data: www.bls.gov | Plot: @ldbailey255") +
  geom_hline(aes(yintercept = 1), lty = 2) +
  geom_text(aes(x = mean(inequality_by_sex$time), y = 0.98, label = "Income parity"), size = 5) +
  geom_text(data = sex_symbol, aes(x = time, y = income_ratio, label = sex_symbol), size = 10) +
  scale_colour_manual(name = "Sex", values = c("#266742", "#3255A2")) +
  scale_x_continuous(breaks = seq(2010, 2021, 1)) +
  scale_y_continuous(limits = c(NA, 1)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_text(vjust = 0.1),
        axis.title.y = element_text(vjust = 0.9),
        axis.text.x = element_text(colour = "black"),
        plot.title = element_text(face = "bold", size = 20),
        plot.subtitle = element_text(size = 10))
  
```

Add custom text and save nice aspect ratio.

```{r, fig.height=14, fig.width=18}

#Load a font from Google Fonts
sysfonts::font_add_google("Oswald", "Oswald", regular.wt = 300)
sysfonts::font_add_google("Ubuntu Mono", "Ubuntu Mono")

#Specify that the showtext package should be used
#for rendering text
showtext::showtext_auto()

ggplot() +
  geom_line(data = inequality_by_sex, aes(x = time, y = income_ratio, group = sex),
            colour = "grey50", size = 0.5, alpha = 0.5) +
  geom_line(data = rolling_mean, aes(x = time, y = rollingmean, colour = sex), size = 1) +
  labs(x = "", y = "<- Less equal                           More equal ->",
       title = "Racial income inequality in the USA",
       subtitle = "Ratio between the income of the lowest and highest earning racial group",
       caption = "Data: www.bls.gov | Plot: @ldbailey255") +
  geom_hline(aes(yintercept = 1), lty = 2) +
  geom_text(aes(x = mean(inequality_by_sex$time), y = 0.98, label = "Income parity"),
            size = 20, family = "Oswald") +
  geom_text(data = sex_symbol, aes(x = time, y = income_ratio, label = sex_symbol), size = 35) +
  scale_colour_manual(name = "Sex", values = c("#266742", "#3255A2")) +
  scale_x_continuous(breaks = seq(2010, 2021, 1)) +
  scale_y_continuous(limits = c(NA, 1)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_text(face = "bold", vjust = 0.9, size = 50, family = "Oswald"),
        axis.text = element_text(colour = "black", size = 30, family = "Oswald"),
        plot.title = element_text(face = "bold", size = 70, family = "Oswald"),
        plot.subtitle = element_text(size = 45, family = "Oswald"),
        plot.caption = element_text(size = 25, family = "Ubuntu Mono"))

ggsave(filename = here::here("./plots/2021/Week9/2021Week9.png"), height = 7, width = 9, dpi = 600, scale = 0.5)

```

