---
title: "Week20_2020"
author: "Liam D. Bailey"
output: html_document
---

```{r, echo = FALSE, include = FALSE}
options(scipen = 200)

library(dplyr)
library(ggplot2)
```

## Goal

The goal in this TidyTuesday is to make an animated plot using nested images. This will necessarily require using Image Magick rather than `gganimate`.

## Preparation

**Read in data**

```{r}

#All TidyTuesday data
tuesdata <- tidytuesdayR::tt_load('2020-05-12')
tuesdata <- tidytuesdayR::tt_load(2020, week = 20)

#Eruption data
eruptions <- tuesdata$eruptions

#Map data
borders <- ggplot2::map_data("world", ".", xlim = NULL, ylim = NULL)

```

**Clean the data**

We can only use data with date information, lat/long, and vei.

There are a number of pre-historic volcanic eruptions in the data, so here we'll just subset those since 1800.

```{r}

eruptions_clean <- eruptions %>% 
  dplyr::filter_at(.vars = vars(starts_with("start"), starts_with("end"), latitude, longitude, vei), ~!is.na(.)) %>% 
  dplyr::filter(start_year >= 1800) %>% 
  dplyr::mutate(start_day = replace(start_day, start_day == 0, 1),
                start_month = replace(start_month, start_month == 0, 1)) %>% 
  dplyr::mutate(fulldate = lubridate::ymd(paste(start_year, start_month, start_day)))

```

## Inset plot: group eruptions by volcano number over time

```{r}

eruptions_cumulative <- eruptions_clean %>% 
  dplyr::group_by(volcano_number, fulldate) %>% 
  dplyr::summarise(daily_eruptions = n()) %>% 
  dplyr::mutate(eruption_cumsum = cumsum(daily_eruptions))

#Work out cutoff for most active volcanoes
eruptions_total <- eruptions_cumulative %>% 
  dplyr::summarise(total_eruptions = sum(daily_eruptions))

big_cutoff <- eruptions_total %>% 
  dplyr::pull(total_eruptions) %>% 
  quantile(0.95)

eruptions_total <- eruptions_total %>% 
  dplyr::mutate(isbig = total_eruptions > big_cutoff) %>% 
  dplyr::select(-total_eruptions)

eruptions_plot_data <- dplyr::left_join(eruptions_cumulative, eruptions_total, by = "volcano_number")

side_plot <- ggplot(data = eruptions_plot_data)+
  geom_line(aes(x = fulldate, y = eruption_cumsum, group = volcano_number, colour = isbig, size = isbig)) +
  scale_colour_manual(values = c("#334E58", "#C52C12")) +
  scale_size_manual(values = c(0.5, 1)) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "#32292F"),
        plot.background = element_rect(fill = "#32292F"),
        axis.title = element_text(colour = "#F8F7FF"),
        axis.text = element_text(colour = "#F8F7FF"),
        axis.line = element_line(colour = "#F8F7FF"),
        axis.ticks = element_line(colour = "#F8F7FF"),
        plot.margin = ggplot2::margin(t = 100, b = 100))

```

## Base map w/ inset

## Create base map

```{r}

base_map <- ggplot() +
  geom_polygon(data = borders, aes(x = long, y = lat, group = group), fill = "#334E58", colour = "#F8F7FF",
               alpha = 0.75, size = 0.5) +
  geom_point(data = eruptions_clean, aes(x = longitude, y = latitude, size = vei), colour = "#C52C12", alpha = 0.75) +
  theme_classic() +
  theme(plot.background = element_rect(fill = "#32292F"),
        panel.background = element_rect(fill = "#32292F"),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")

base_map

```

## Animate base map

```{r}

ggplot() +
  geom_polygon(data = borders, aes(x = long, y = lat, group = group), fill = "#334E58", colour = "#F8F7FF",
               alpha = 0.75, size = 0.5) +
  geom_point(data = eruptions_clean, aes(x = longitude, y = latitude, size = vei), colour = "#C52C12", alpha = 0.75) +
  theme_classic() +
  theme(plot.background = element_rect(fill = "#32292F"),
        panel.background = element_rect(fill = "#32292F"),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")+
  #Start gganimate code
  gganimate::transition_time(time = fulldate)+
  labs(title = "Year: {frame_time}")+
  gganimate::shadow_mark(alpha = 0.25, wrap = FALSE, size = 2, exclude_layer = 2:3)+
  gganimate::ease_aes("linear")

```

## Create combo plot using patchwork

```{r}

library(patchwork)

base_map + side_plot + 
  plot_layout(widths = c(2, 1))

```


```{r}

ggsave(filename = "../plots/map_test.jpeg", dpi = 600, width = 12, height = 7)

```

```{r}

a <- ggplot(means_df, aes(x = IV,y = means, fill = IV)) +
  geom_bar(stat = "identity") +
  geom_point(aes(x = IV, y = DV), data = df, alpha = .25) +
  geom_errorbar(aes(ymin = means - sem, ymax = means + sem), width = .2) +
  theme_classic() +
  transition_states(
    states = sims,
    transition_length = 2,
    state_length = 1
  ) + 
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')
  
a_gif <- animate(a, width = 240, height = 240)

b <- ggplot(stats_df, aes(x = ts))+
  geom_vline(aes(xintercept = ts, frame = sims))+
  geom_line(aes(x=x,y=y),
            data = data.frame(x = seq(-5,5, .1),
                              y = dt(seq(-5,5, .1), df = 18))) +
  theme_classic() +
  ylab("density") +
  xlab("t value") +
  transition_states(
    states = sims,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')

b_gif <- animate(b, width = 240, height = 240)

a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1]))
for(i in 2:100){
  combined <- image_append(c(a_mgif[i], b_mgif[i]))
  new_gif <- c(new_gif, combined)
}

new_gif

```

