---
title: "Week25_2022"
author: "Liam D. Bailey"
date: '2022-07-05'
output: html_document
---

```{r, include = FALSE}

#Options

```

## Load data

```{r}

juneteenth_data <- tidytuesdayR::tt_load(2022, week = 25)

```

## Load required packages

```{r, message=FALSE}

library(ggplot2) #For plotting
library(dplyr) #For data wrangling
library(patchwork) #To combine ggplot objects
library(showtext) #For custom fonts
library(gt) #For creating nice tables
library(scales) #For creating nice palettes
library(lubridate) #For working with dates
library(ggtext) #For richtext
library(here) #For easier relative paths
library(skimr) #For exploring data

showtext::showtext_auto()
sysfonts::font_add_google(name = "Quicksand", family = "Quicksand")
sysfonts::font_add_google(name = "Source Code Pro", family = "Source_Code_Pro")
sysfonts::font_add_google(name = "Cherry Swash", family = "Cherry_Swash")
sysfonts::font_add_google(name = "Alfa Slab One", family = "Alfa_Slab_One")
sysfonts::font_add_google(name = "Cutive Mono", family = "Cutive_Mono")

### NOTE ###
#The below fonts are custom fonts installed on my computer locally
#They will not work on your system by default!

#If you want to use them in the plot you'll need to download the .otf files 
#and add them into the font search path where sysfonts can find them.
#Find your current font search paths (and add new paths) using `sysfonts::font_paths()`

#Download here https://www.kontrapunkt.com/typetester
font_add(family = "Miki",
         regular = "KontrapunktMiki-Regular.otf",
         bold = "KontrapunktMiki-Bold.otf")
#Download trial font here https://www.vocaltype.co/buy-du-bois
font_add(family = "DuBois",
         regular = "VTCDuBoisTrial-Regular.otf",
         bold = "VTCDuBoisTrial-BoldWide.otf")

```

## Explore data

### Census data

```{r}
census <- juneteenth_data$census
skim(census)
```

- We have data from 1790 - 1870.
- Have data on white, black(enslaved), and black(free) in all years
- Region is always available, but 'division' is sometimes missing. What is 'division'?

```{r}
unique(census$division) #Smaller breakdown of regions
```

I think that we have totals for the whole of US, each region and each sub-division.

## Plot data

### Census data

#### Plot idea 1: Combined bar graph showing proportion of slaves v. free in Union v. Confederacy

```{r}
subset_data <- census %>% 
  #Remove US total and West (we want to compare difference b/w confederate and non-confederate)
  #Don't worry about breaking down by division, so keep this as NA
  filter(!region %in% c("USA Total", "West") & is.na(division)) %>% 
  #Group Northeast and Midwest as 'Northern States'
  #Keep 'South' as is
  #This is not *exactly* right, because some northern states are in the 'South' region (e.g. Maryland, Delaware)
  mutate(civil_war_state = case_when(region %in% c("Northeast", "Midwest") ~ "Union",
                                     region == "South" ~ "South")) %>% 
  #group and determine number of black, free black, enslaved black
  group_by(civil_war_state, year) %>% 
  summarise(black = sum(black),
            black_free = sum(black_free),
            black_enslaved = sum(black_slaves), .groups = "drop") %>% 
  #Percentage for each group
  mutate(free_perc = (black_free/black)*100,
         enslaved_perc = (black_enslaved/black)*100)
```

```{r}

library(ggnewscale)
ggplot(data = subset_data) +
  #Create black and white rectangles behind plot
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = 0, ymax = -Inf), fill = "grey10") +
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = 0, ymax = Inf), fill = "grey90") +
  #Add enslaved data
  geom_col(aes(x = year, y = -enslaved_perc,
               group = civil_war_state,
               fill = civil_war_state),
           position = position_dodge(width = 8),
           colour = "grey10",
           width = 7) +
  scale_fill_manual(values = c("#610519", "#002766")) +
  # geom_text(aes(x = year, y = -enslaved_perc - 3,
  #               label = paste0(round(enslaved_perc), "%"),
  #               group = civil_war_state),
  #           colour = "white", position = position_dodge(width = 8),
  #           size = 3) +
  #Add free data
  new_scale_fill() +
  geom_col(aes(x = year, y = free_perc,
               group = civil_war_state,
               fill = civil_war_state),
           position = position_dodge(width = 8),
           colour = "grey10",
           width = 7) +
  scale_fill_manual(values = c("#910825", "#003b99")) +
  # geom_text(aes(x = year, y = free_perc + 3,
  #               label = paste0(round(free_perc), "%"),
  #               group = civil_war_state),
  #           colour = "black", position = position_dodge(width = 8),
  #           size = 3) +
  #Add horizontal line at 0
  geom_hline(yintercept = 0, colour = "grey10", size = 0.5) +
  #Adjust scales
  scale_x_continuous(breaks = seq(1790, 1870, 10), position = "top",
                     limits = c(1785, 1875)) +
  scale_y_continuous(breaks = c(-100, -50, 0, 50, 100),
                     labels = c("100%", "50%", "0%", "50%", "100%")) +
  #Theme adjustment
  theme_classic() +
  theme(axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 15, colour = "grey10"),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 12, colour = "grey10"),
        legend.position = "none",
        plot.background = element_rect(fill = "#e7dbc0"))
```

It's fine, but not super convincing. Also grouping NE and Midwest together doesn't give a complete picture.

#### Plot idea 2: For each region with state legend

```{r}
subset_data <- census %>% 
  #Remove US total and West (we want to compare difference b/w confederate and non-confederate)
  #Don't worry about breaking down by division, so keep this as NA
  filter(!region %in% c("USA Total", "West") & is.na(division)) %>% 
  #group and determine number of black, free black, enslaved black per region
  group_by(region, year) %>% 
  summarise(black = sum(black),
            black_free = sum(black_free),
            black_enslaved = sum(black_slaves), .groups = "drop") %>% 
  #Percentage for each group
  mutate(free_perc = (black_free/black)*100,
         enslaved_perc = (black_enslaved/black)*100)
```

```{r}
ggplot(data = subset_data) +
  #Create black and white rectangles behind plot
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = 0, ymax = -Inf), fill = "grey10") +
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = 0, ymax = Inf), fill = "grey90") +
  #Add enslaved data
  geom_col(aes(x = year, y = -enslaved_perc,
               group = region,
               fill = region),
           position = position_dodge(width = 8),
           colour = "grey10",
           width = 7) +
  geom_text(aes(x = year, y = -enslaved_perc - 3,
                label = paste0(round(enslaved_perc), "%"),
                group = region),
            colour = "white", position = position_dodge(width = 8),
            size = 3) +
  #Add free data
  geom_col(aes(x = year, y = free_perc,
               group = region,
               fill = region),
           position = position_dodge(width = 8),
           colour = "grey10",
           width = 7) +
  geom_text(aes(x = year, y = free_perc + 3,
                label = paste0(round(free_perc), "%"),
                group = region),
            colour = "black", position = position_dodge(width = 8),
            size = 3) +
  #Add horizontal line at 0
  geom_hline(yintercept = 0, colour = "grey10", size = 0.5) +
  #Adjust scales
  scale_x_continuous(breaks = seq(1790, 1870, 10), position = "top",
                     limits = c(1785, 1875)) +
  scale_y_continuous(breaks = c(-100, -50, 0, 50, 100),
                     labels = c("100%", "50%", "0%", "50%", "100%")) +
  #Facet
  facet_wrap(facets = ~region, ncol = 1) +
  #Theme adjustment
  theme_classic() +
  theme(axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 15, colour = "grey10"),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 12, colour = "grey10"),
        legend.position = "none",
        plot.background = element_rect(fill = "#e7dbc0"))
```

Try same plots but individual per region and flipped

```{r}
plot_data <- subset_data %>% 
  filter(region == "Midwest")

midwest_bars <- ggplot() +
  #Create black and white rectangles behind plot
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = -Inf), fill = "grey10") +
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = Inf), fill = "grey90") +
  #Add enslaved data
  geom_col(data = plot_data,
           aes(x = year, y = -enslaved_perc),
           position = position_dodge(width = 8),
           fill = "grey10",
           colour = "grey10",
           width = 7) +
  geom_richtext(data = plot_data,
            aes(x = year, y = -enslaved_perc - 7,
                label = paste0(round(enslaved_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 1,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add free data
  geom_col(data = plot_data,
           aes(x = year, y = free_perc),
           position = position_dodge(width = 8),
           fill = "grey50",
           colour = "grey50",
           width = 7) +
  geom_richtext(data = plot_data,
            aes(x = year, y = free_perc + 7,
                label = paste0(round(free_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 0,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add additional text specifying sections
  geom_text(aes(x = c(1785),
                y = c(-5),
                label = c("ENSLAVED")),
            colour = "grey10", hjust = 1,
            family = "DuBois", fontface = "bold",
            size = 3) +
  geom_text(aes(x = c(1785),
                y = c(5),
                label = c("FREE")),
            colour = "grey50", hjust = 0,
            family = "DuBois", fontface = "bold",
            size = 3) +
  #Add horizontal line at 0
  geom_hline(yintercept = 0, colour = "grey10", size = 0.5) +
  #Adjust scales
  scale_x_reverse(breaks = seq(1790, 1870, 10),
                  limits = rev(c(1785, 1875))) +
  scale_y_continuous(breaks = c(-100, -50, 0, 50, 100),
                     labels = paste0(c("100", "50", "0", "50", "100"),
                                     "<span style='font-family:Cutive_Mono'>%</span>"),
                     limits = c(-110, 110)) +
  #Flip
  coord_flip(clip = "off") +
  #Theme adjustment
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_markdown(size = 10, colour = "grey10",
                                   family = "DuBois"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none",
        plot.background = element_blank(),
        panel.background = element_blank())

midwest_bars

```

```{r}
plot_data <- subset_data %>% 
  filter(region == "Northeast")

NE_bars <- ggplot(data = plot_data) +
  #Create black and white rectangles behind plot
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = -Inf), fill = "grey10") +
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = Inf), fill = "grey90") +
  #Add enslaved data
  geom_col(aes(x = year, y = -enslaved_perc),
           position = position_dodge(width = 8),
           fill = "grey10",
           colour = "grey10",
           width = 7) +
  geom_richtext(aes(x = year, y = -enslaved_perc - 7,
                label = paste0(round(enslaved_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 1,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add free data
  geom_col(aes(x = year, y = free_perc),
           position = position_dodge(width = 8),
           fill = "grey50",
           colour = "grey50",
           width = 7) +
  geom_richtext(aes(x = year, y = free_perc + 7,
                label = paste0(round(free_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 0,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add horizontal line at 0
  geom_hline(yintercept = 0, colour = "grey10", size = 0.5) +
  #Adjust scales
  scale_x_reverse(breaks = seq(1790, 1870, 10),
                  limits = rev(c(1785, 1875))) +
  scale_y_continuous(breaks = c(-100, -50, 0, 50, 100),
                     labels = paste0(c("100", "50", "0", "50", "100"),
                                     "<span style='font-family:Cutive_Mono'>%</span>"),
                     limits = c(-110, 110)) +
  #Flip
  coord_flip(clip = "off") +
  #Theme adjustment
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_markdown(size = 10, colour = "grey10",
                                   family = "DuBois"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none",
        plot.background = element_blank(),
        panel.background = element_blank())

NE_bars

```

```{r}
plot_data <- subset_data %>% 
  filter(region == "South")

South_bars <- ggplot(data = plot_data) +
  #Create black and white rectangles behind plot
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = -Inf), fill = "grey10") +
  # geom_rect(aes(xmin = -Inf, xmax = Inf,
  #               ymin = 0, ymax = Inf), fill = "grey90") +
  #Add enslaved data
  geom_col(aes(x = year, y = -enslaved_perc),
           position = position_dodge(width = 8),
           fill = "grey10",
           colour = "grey10",
           width = 7) +
  geom_richtext(aes(x = year, y = -enslaved_perc - 7,
                label = paste0(round(enslaved_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 1,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add free data
  geom_col(aes(x = year, y = free_perc),
           position = position_dodge(width = 8),
           fill = "grey50",
           colour = "grey50",
           width = 7) +
  geom_richtext(aes(x = year, y = free_perc + 7,
                label = paste0(round(free_perc), "<span style='font-family:Cutive_Mono'>%</span>")),
            colour = "black",
            size = 2.5, hjust = 0,
            family = "DuBois",
            label.colour = NA, fill = NA) +
  #Add horizontal line at 0
  geom_hline(yintercept = 0, colour = "grey10", size = 0.5) +
  #Adjust scales
  scale_x_reverse(breaks = seq(1790, 1870, 10),
                  limits = rev(c(1785, 1875))) +
  scale_y_continuous(breaks = c(-100, -50, 0, 50, 100),
                     labels = paste0(c("100", "50", "0", "50", "100"),
                                     "<span style='font-family:Cutive_Mono'>%</span>"),
                     limits = c(-110, 110)) +
  #Flip
  coord_flip(clip = "off") +
  #Theme adjustment
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_markdown(size = 10, colour = "grey10",
                                   family = "DuBois"),
        axis.title.x = element_blank(),
        axis.text.x = element_markdown(size = 10, colour = "grey10",
                                   family = "DuBois"),
        axis.line.x = element_blank(),
        legend.position = "none",
        plot.background = element_blank(),
        panel.background = element_blank())

South_bars

```

Looks ok. Create separate maps for each of the regions using polygons we got last time.

```{r, cache=TRUE}
state <- raster::getData("GADM", country = "USA", level = 1)
state_sf <- sf::st_as_sf(state)

slave_states <- data.frame(state = c("Kansas", "Missouri", "Iowa", "Minnesota",
                                     "Wisconsin", "Illinois", "Indiana",
                                     "Michigan", "Ohio", "Pennsylvania", "New York", "Vermont",
                                     "Connecticut", "New Jersey", "Rhode Island",
                                     "Maine",
                                     "New Hampshire", "Massachusetts", "Texas", "Arkansas",
                                     "Louisiana",
                                     "Mississippi", "Alabama", "Tennessee", "Kentucky",
                                     "Georgia", "Florida", "South Carolina",
                                     "North Carolina", "Virginia", "West Virginia",
                                     "Maryland", "Delaware", "District of Columbia"),
                           side = c("Free", "Slave", "Free", "Free",
                                    "Free", "Free", "Free",
                                    "Free", "Free", "Free", "Free", "Free",
                                    "Free", "Free", "Free",
                                    "Free",
                                    "Free", "Free", "Slave", "Slave",
                                    "Slave", "Slave", "Slave", "Slave",
                                    "Slave", "Slave", "Slave",
                                    "Slave", "Slave", "Slave", "Slave",
                                    "Slave", "Slave", NA))

#Separate into our different regions
#For midwest, we exclude states that didn't exist during civil war
#This includes: Dakotas, Nebraska (made a state in 1867)
midwest <- state_sf %>% 
  filter(NAME_1 %in% c("Kansas", "Missouri", "Iowa", "Minnesota",
                       "Wisconsin", "Illinois", "Indiana",
                       "Michigan", "Ohio")) %>% 
  left_join(slave_states, by = c("NAME_1" = "state")) %>% 
  #simplify down to make it easier to plot
  sf::st_simplify(dTolerance = 1e3)

northeast <- state_sf %>% 
  filter(NAME_1 %in% c("Pennsylvania", "New York", "Vermont",
                       "Connecticut", "New Jersey", "Rhode Island",
                       "Maine",
                       "New Hampshire", "Massachusetts")) %>% 
  left_join(slave_states, by = c("NAME_1" = "state")) %>% 
  #simplify down to make it easier to plot
  sf::st_simplify(dTolerance = 1e3)

#Remove Oklahoma (not state yet)
south <- state_sf %>% 
  filter(NAME_1 %in% c("Texas", "Arkansas", "Louisiana",
                       "Mississippi", "Alabama", "Tennessee", "Kentucky",
                       "Georgia", "Florida", "South Carolina",
                       "North Carolina", "Virginia", "West Virginia",
                       "Maryland", "Delaware", "District of Columbia")) %>% 
  left_join(slave_states, by = c("NAME_1" = "state")) %>% 
  #simplify down to make it easier to plot
  sf::st_simplify(dTolerance = 1e3)

midwest_map <- ggplot() +
  geom_sf(data = midwest,
          aes(fill = side),
          colour = "grey60", size = 0.25) +
  scale_fill_manual(values = c("#003b99", "#920725")) +
  theme_void() +
  theme(legend.position = "none")

northeast_map <- ggplot() +
  geom_sf(data = northeast, 
          aes(fill = side),
          colour = "grey60", size = 0.25) +
  scale_fill_manual(values = c("#003b99")) +
  theme_void() +
  theme(legend.position = "none")

south_map <- ggplot() +
  geom_sf(data = south, 
          aes(fill = side),
          colour = "grey60", size = 0.25) +
  scale_fill_manual(values = c("#920725")) +
  theme_void() +
  theme(legend.position = "none")

```

Create a timeline to put at the bottom

```{r}
timeline <- ggplot() +
  geom_segment(aes(x = as.Date("1600-01-01"), xend = as.Date("1870-01-01"), y = 0.5, yend = 0.5),
               size = 2, colour = "grey10") +
  #Independence
  geom_segment(aes(x = as.Date("1776-07-04"), xend = as.Date("1776-07-04"),
                   y = 0.5, yend = 0.8)) +
  geom_richtext(aes(x = as.Date("1776-07-04"),
                    y = 0.75,
                    label = "<span style='font-size:15px'>**July 4, 1776**</span><br>Declaration of independence"),
                hjust = 0, family = "DuBois",
                label.colour = NA, fill = NA) +
  #Abolish the slave trade
  geom_segment(aes(x = as.Date("1808-01-01"), xend = as.Date("1808-01-01"),
                   y = 0.5, yend = 0.2)) +
  geom_richtext(aes(x = as.Date("1808-01-01"),
                    y = 0.275,
                    label = "<span style='font-size:15px'>**January 1, 1808**</span><br>Atlantic slave trade abolished<br>*Domestic* slave trade continues"),
                hjust = 1, family = "DuBois",
                label.colour = NA, fill = NA) +
  #Confederacy established
  geom_segment(aes(x = as.Date("1861-01-01"), xend = as.Date("1861-01-01"),
                   y = 0.5, yend = 0.2)) +
  geom_richtext(aes(x = as.Date("1861-01-01"),
                    y = 0.275,
                    label = "<span style='font-size:15px'>**February, 1861**</span><br>Confederate slave states<br>break from the Union"),
                hjust = 1, family = "DuBois",
                label.colour = NA, fill = NA) +
  #Emancipation
  geom_segment(aes(x = as.Date("1865-12-01"), xend = as.Date("1865-12-01"),
                   y = 0.5, yend = 0.8)) +
  geom_richtext(aes(x = as.Date("1865-12-01"),
                    y = 0.725,
                    label = "<span style='font-size:15px'>**December, 1865**</span><br>The 13th Amendment<br>abolishes slavery"),
                hjust = 1, family = "DuBois",
                label.colour = NA, fill = NA) +
  scale_y_continuous(limits = c(0.1, 0.9)) +
  scale_x_date(breaks = seq(as.Date("1600-01-01"), as.Date("1870-01-01"), "10 years"),
               date_labels = "%Y") +
  coord_equal(ratio = 12000) +
  labs(title = "TIMELINE OF AMERICAN SLAVERY") +
  theme_void() +
  theme(panel.background = element_blank(),
        plot.background = element_blank(),
        axis.text.x = element_text(size = 10, colour = "grey10",
                                   family = "DuBois"),
        panel.grid.major.x = element_line(colour = "grey90"),
        plot.title = element_text(size = 15, colour = "black", family = "DuBois", face = "bold"))
```

Combine altogether as grobs

```{r, warning = FALSE}

midwest_bars_grob <- ggplotGrob(midwest_bars)
midwest_map_grob <- ggplotGrob(midwest_map)

south_bars_grob <- ggplotGrob(South_bars)
south_map_grob <- ggplotGrob(south_map)

northeast_bars_grob <- ggplotGrob(NE_bars)
northeast_map_grob <- ggplotGrob(northeast_map)

timeline_grob <- ggplotGrob(timeline)

```

Make complete plot

```{r}

ggplot() +
  scale_x_continuous(limits = c(0, 2.25)) +
  scale_y_continuous(limits = c(-0.1, 1.1)) +
  coord_fixed(ratio = 1, clip = "off") +
  #Midest
  annotation_custom(grob = midwest_map_grob,
                    xmin = 0, xmax = 0.5,
                    ymin = 0.6, ymax = 1) +
  annotation_custom(grob = midwest_bars_grob,
                    xmin = 1.2, xmax = 1.8,
                    ymin = 0.7, ymax = 1.1) +
  #NE
  annotation_custom(grob = northeast_map_grob,
                    xmin = 0.6, xmax = 1.1,
                    ymin = 0.55, ymax = 0.95) +
  annotation_custom(grob = northeast_bars_grob,
                    xmin = 1.2, xmax = 1.8,
                    ymin = 0.3, ymax = 0.7) +
  #South
  annotation_custom(grob = south_map_grob,
                    xmin = 0, xmax = 0.9,
                    ymin = -0.2, ymax = 0.7) +
  annotation_custom(grob = south_bars_grob,
                    xmin = 1.2, xmax = 1.8,
                    ymin = -0.1, ymax = 0.3) +
  #Add lines and label midwest
  geom_segment(aes(y = 0.8, yend = 0.95,
                   x = 0.25, xend = 0.25),
               size = 1.5, colour = "grey10",
               linejoin = "round", lineend = "round") +
  geom_segment(aes(y = 0.95, yend = 0.95,
                   x = 0.25, xend = 1.2),
               size = 1.5, colour = "grey10",
               linejoin = "round") +
  geom_text(aes(y = 0.975, x = 1.2, label = "MIDWEST"),
            size = 5, hjust = 1, family = "DuBois", fontface = "bold") +
  geom_point(aes(x = 0.25, y = 0.8),
             shape = 21, colour = "black",
             fill = "white", stroke = 2,
             size = 5) +
  #Add lines NE
  geom_segment(aes(y = 0.75, yend = 0.55,
                   x = 0.9, xend = 0.9),
               size = 1.5, colour = "grey10",
               linejoin = "round", lineend = "round") +
  geom_segment(aes(y = 0.55, yend = 0.55,
                   x = 0.9, xend = 1.2),
               size = 1.5, colour = "grey10",
               linejoin = "round") +
  geom_text(aes(y = 0.525, x = 1.2, label = "NORTHEAST"),
            size = 5, hjust = 1, family = "DuBois", fontface = "bold") +
  geom_point(aes(x = 0.9, y = 0.75),
             shape = 21, colour = "black",
             fill = "white", stroke = 2,
             size = 5) +
  #Add lines South
  geom_segment(aes(y = 0.25, yend = 0.1,
                   x = 0.45, xend = 0.45),
               size = 1.5, colour = "grey10",
               linejoin = "round", lineend = "round") +
  geom_segment(aes(y = 0.1, yend = 0.1,
                   x = 0.45, xend = 1.2),
               size = 1.5, colour = "grey10",
               linejoin = "round") +
  geom_text(aes(y = 0.075, x = 1.2, label = "SOUTH"),
            size = 5, hjust = 1, family = "DuBois", fontface = "bold") +
  geom_point(aes(x = 0.45, y = 0.25),
             shape = 21, colour = "black",
             fill = "white", stroke = 2,
             size = 5) +
  #Explain slave v. non-slave states
  geom_segment(aes(y = 0.675, yend = 0.51,
                   x = 0.1, xend = 0.1),
               size = 0.75, colour = "grey10") +
  geom_text(aes(x = 0.09, y = 0.55,
                label = "Free\nstates"),
            hjust = 1, colour = "#003b99",
            family = "DuBois", fontface = "bold") +
  geom_segment(aes(y = 0.675, yend = 0.51,
                   x = 0.2, xend = 0.2),
               size = 0.75, colour = "grey10") +
  geom_text(aes(x = 0.21, y = 0.55,
                label = "Slave\nstates"),
            hjust = 0, colour = "#920725",
            family = "DuBois", fontface = "bold") +
  #Add info about each region
  geom_richtext(aes(x = 1.85, y = 0.5,
                    label = "The Northeast <span style='color:#003b99;font-size:15px'>**Free states**</span><br>had almost no enslaved Black<br>population by 1830"),
                size = 3.5, hjust = 0,
                label.colour = NA, fill = NA, family = "DuBois") +
  geom_richtext(aes(x = 1.85, y = 0.15,
                    label = "Southern <span style='color:#920725;font-size:15px'>**Slave states**</span><br>had almost no free Black<br>population until after the Civil War"),
                size = 3.5, hjust = 0,
                label.colour = NA, fill = NA, family = "DuBois") +
  geom_richtext(aes(x = 1.85, y = 0.9,
                    label = "With one <span style='color:#920725;font-size:15px'>**Slave state**</span><span style='font-family:Cutive_Mono'>(</span>Missouri<span style='font-family:Cutive_Mono'>)</span><br>the Midwest still had<br>a substaintial enslaved<br>population before 1870"),
                size = 3.5, hjust = 0,
                label.colour = NA, fill = NA, family = "DuBois") +
  #Add vertical line across all plots
  geom_segment(aes(x = 1.543, xend = 1.543,
                   y = -0.05, yend = 1.0825),
               size = 0.5, colour = "grey10") +
  #Titles
  labs(title = "GEOGRAPHY OF AMERICAN SLAVERY",
       subtitle = "Slavery was prolific in the South before it was abolished in 1865",
       caption = "Data: US Census Archives | Plot: @ldbailey255<br>Geographic groups represent US Census regions<br>Nebraska, Oklahoma, N. Dakota and S. Dakota were not states during this period so were excluded from the maps") +
  #Theme
  theme_void() +
  theme(panel.background = element_rect(fill = "#e7dbc0", colour = NA),
        plot.background = element_rect(fill = "#e7dbc0", colour = NA),
        plot.title = element_text(size = 25, colour = "black", family = "DuBois", face = "bold"),
        plot.subtitle = element_markdown(size = 18, colour = "black", family = "DuBois"),
        plot.caption = element_markdown(size = 9, colour = "black", family = "Cutive_Mono"))

ggsave(filename = here("./plots/2022/Week25.png"), width = 12, height = 7, dpi = 600)

```