---
title: "Week47_2019_Nov19th"
output: html_document
---

Set options and load packages for plotting

```{r}

#Change how R deals with scientific notation
options(scipen = 100)

#Load required libraries
library(ggplot2)
library(egg)
library(extrafont)
library(dplyr)
library(grid)
library(png)
library(ggtext)
library(ggraph)

devtools::load_all()

```

Load data

```{r}

nz_bird <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-19/nz_bird.csv")

```

We have two ideas:

1. Show the top two birds and the distribution of their votes that came from preference flow from other birds.

2. Network diagram to show how voter preferences combine.

3. Create an animation to how the real time spread of votes over time.

# Idea 1:

Format the data for our purposes.

- Create a unique value for every group of 5 rows (i.e. a unique vote ID)

```{r}

nr_unique <- length(which(nz_bird$vote_rank == "vote_5"))
voterID  <- rep(1:nr_unique, 5)
voterID  <- voterID[order(voterID)]

nz_bird <- nz_bird %>% 
  dplyr::mutate(voterID = {{voterID}},
                bird_breed = iconv(bird_breed, "UTF-8")) %>% 
  dplyr::filter(!is.na(bird_breed))

#For every unique voter, create a list of their votes
#and assign their first vote and current vote (which will identical before runoff)
nz_birds_processed <- nz_bird %>% 
  dplyr::group_by(voterID) %>% 
  dplyr::summarise(votes = list(bird_breed),
                   first_vote = first(bird_breed),
                   current_vote = first(bird_breed))

```

Now we want to create a function that will carry out the instant runoff

```{r}

#end_grps is the number of groups that will be left over
#after the runoff. It is 2 by default, but we may want to do fewer if we want to animate the process
runoff <- function(end_grps = 2, data){
  
  #Group votes by each bird (excluding NAs)
  bird_grps <- data %>%
    dplyr::filter(!is.na(current_vote)) %>%
    dplyr::group_by(current_vote) %>% 
    dplyr::summarise(total_votes = n()) %>% 
    dplyr::arrange(-total_votes)
  
  #Create a list to store eliminated birds
  eliminated <- list()
  
  pb <- dplyr::progress_estimated(n = nrow(bird_grps) - 1)
  
  #While there are more than 2 birds left...
  while(nrow(bird_grps) > end_grps){
    
    pb$print()$tick()
    
    #Determine the bird with the lowest votes currently
    lowest_votes = bird_grps %>% 
      dplyr::filter(!is.na(current_vote)) %>% 
      dplyr::slice(n()) %>% 
      dplyr::pull(current_vote)
    
    #Add this bird to the eliminated list
    eliminated <- append(eliminated, lowest_votes)
    
    #For each row check if they are currently voting for an eliminated bird
    data <- data %>% 
      dplyr::mutate(current_vote = purrr::map2_chr(.x = data$votes, .y = data$current_vote,
                                               .f = function(all_votes, current_vote, lowest_votes){
    
    #If bird has been eliminated, move to the next preference that is not eliminated
    #If the current vote is 'NA' this mean the 5 votes have been exhausted
    while(current_vote %in% eliminated){
      
      current_vote <- all_votes[which(all_votes == current_vote) + 1]
      
    }
    
    return(current_vote) 
    
}, lowest_votes))
    
    #Regroup data by birds and remove NAs (i.e. exhausted votes)
    bird_grps <- data %>% 
      dplyr::filter(!is.na(current_vote)) %>% 
      dplyr::group_by(current_vote) %>%
      dplyr::summarise(total_votes = n()) %>%
      dplyr::arrange(-total_votes)
  
  }
  
  return(data %>% 
           dplyr::filter(!is.na(current_vote)))
  
}

```

Carry out instant run-off to find the top two birds

```{r}

top2_runoff <- runoff(data = nz_birds_processed)

```

Identify the top voting groups and classify the rest as other

```{r}

#Determine the top 5 birds for primary votes
top5 <- top2_runoff %>% 
  dplyr::group_by(first_vote) %>% 
  dplyr::summarise(n = n()) %>% 
  dplyr::arrange(-n) %>% 
  dplyr::slice(1:5)

#For all birds not in the top5 classify them as 'other'
top2_runoff <- top2_runoff %>% 
  dplyr::mutate(fill_grp = factor(purrr::map_chr(.x = first_vote, .f = ~{
    
    if(..1 %in% top5$first_vote){
      
      return(..1)
      
    } else {
      
      return("Other")
      
    }
    
    #Relevel to fit our plot
  }), levels = c("Kakapo", "Yellow-eyed penguin", "Kea", "Black Robin", "Banded Dotterel", "Other")))

#Read in bird png images
other   <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "other.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))
kakapo   <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "kakapo.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))
kea      <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "kea.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))
dotterel <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "dotterel.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))
robin    <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "blackrobin.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))
penguin  <- grid::rasterGrob(png::readPNG(source = system.file("extdata/NZ_birds", "penguin.png", package = "TidyTuesday", mustWork = TRUE)),
                                            width = unit(2.5, "cm"), height = unit(2.5, "cm"))

#Add png images with locations and text to a tibble
bird_legend <- tibble(x = 1,
                      y = seq(0.15, 0.85, length.out = 6)) %>%
  mutate(grob = list(other, kakapo, kea, dotterel, robin, penguin),
         text = c("<span style = 'color:#6a3b7d'>Other</span>",
                  "<span style = 'color:#98b02c'>Kakapo</span>",
                  "<span style = 'color:#4b5439'>Kea</span>",
                  "<span style = 'color:#b5773a'>Banded <br> Dotterel</span>",
                  "<span style = 'color:black'>Black Robin</span>",
                  "<span style = 'color:#4b4c61'>Yellow-eyed <br> Penguin</span>"))

#Create bar graph and convert to grob
bars <- ggplot() +
  geom_bar(data = top2_runoff, aes(x = current_vote, fill = fill_grp), colour = "black", size = 1) +
  scale_y_continuous(limits = c(0, 11000), breaks = seq(0, 10000, 2500), name = "Total votes") +
  scale_fill_manual(values = c("#98b02c", "#4b4c61", "#4b5439", "black", "#b5773a", "#6a3b7d")) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.y = element_text(colour = "black", family = "Playfair Display Black", size = 15),
        axis.text.x = element_text(colour = "black", family = "Playfair Display Black", size = 18),
        axis.title = element_text(colour = "black", family = "Playfair Display Black", size = 20))
  
bars <- ggplot2::ggplotGrob(bars)

#Create a blank canvas and add the bar graph and png legend
ggplot()+
  scale_x_continuous(limits = c(0, 1.25), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  labs(title = "Who is New Zealand's favourite bird?",
       subtitle = "The <span style = color:#4b4c61> Yellow-eyed Penguin </span> (or Hoiho) beat the  <span style = color:#98b02c> Kakapo </span> to be voted NZ Bird of the Year.<br>
       Colours show the proportion of first-choice votes from other species <br> assigned to the two finalists following the instant run-off.",
       caption = "Plot: @ldbailey255 | Data: dragonfly.co.nz") +
  annotation_custom(grob = bars, xmin = 0, xmax = 0.95, ymin = 0, ymax = 1) +
  egg::geom_custom(data = bird_legend, aes(x = x, y = y, data = grob), grob_fun = "identity") +
  ggtext::geom_richtext(data = bird_legend, aes(x = x + 0.15, y = y, label = text), size = 7, family = "Playfair Display Black") +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, family = "Playfair Display Black", size = 30),
        plot.subtitle = element_markdown(hjust = 0.5, family = "Playfair Display Black", size = 20),
        plot.caption = element_text(family = "Playfair Display Black", size = 12))

#Save
ggplot2::ggsave(filename = "../plots/Week47_2019.png", height = 10, width = 12, dpi = 300, type = "cairo")

```

# Idea 2: Network diagram

First we need to convert our data to be in the correct format, where we should have a 'from' and 'to' column. Link only first votes with all additional votes.

```{r}

nz_bird_network <- nz_bird %>% 
  dplyr::mutate(bird_grp = dplyr::case_when(
                  grepl(pattern = "Penguin|penguin", x = bird_breed) ~ "Penguins",
                  grepl(pattern = "Kiwi", x = bird_breed) ~ "Kiwis",
                  grepl(pattern = "Albatross|Petrel|Shearwater|Gannet|Skua|Shag|Gull|Tern|Mollymawk", x = bird_breed) ~ "Seabirds",
                  grepl(pattern = "Dotterel|Stilt|Plover|Godwit|Oystercatcher|Wrybill", x = bird_breed) ~ "Shorebirds",
                  grepl(pattern = "Kākā|Kākāpō|Kea|Parakeet", x = bird_breed) ~ "Parrots",
                  grepl(pattern = "Harrier|Falcon|Owl|Morepork", x = bird_breed) ~ "Birds of Prey",
                  grepl(pattern = "Crake|Spoonbill|Heron|Bittern|Takahē|Weka|Pūkeko", x = bird_breed) ~ "Waders",
                  grepl(pattern = "Robin|Silvereye|Wren|Fantail|Warbler|Saddelback|Kōkako|Tūī|Bellbird|Tomtit|Mōhua|Rifleman|Stitchbird|Swallow|Fernbird|", x = bird_breed) ~ "Passerines",
                  grepl(pattern = "Teal|Duck|Dabchick", x = bird_breed) ~ "Water birds",
                  grepl(pattern = "Kererū", x = bird_breed) ~ "Pigeons",
                  grepl(pattern = "Cuckoo", x = bird_breed) ~ "Cuckoos")) %>%
  dplyr::group_by(voterID) %>% 
  dplyr::arrange(vote_rank, .by_group = TRUE) %>% 
  dplyr::mutate(from = first(bird_breed), from_grp = first(bird_grp),
                to = bird_breed, to_grp = bird_grp,
                withingrp = from_grp == to_grp) %>% 
  dplyr::filter(!is.na(to) & from != to) %>% 
  dplyr::select(from, to, withingrp, voterID) %>% 
  dplyr::ungroup()

```

Test graph

```{r}

test_data <- nz_bird_network %>% 
  dplyr::filter(voterID < 100)

graph <- tidygraph::as_tbl_graph(test_data) %>% 
  #Remove cases where there are redundant edges (i.e. the same link occurs multiple times)
  tidygraph::activate(edges) %>% 
  filter(!tidygraph::edge_is_multiple()) %>% 
  #Order nodes to minimise distances
  tidygraph::activate(nodes) %>% 
  dplyr::mutate(n_rank = tidygraph::node_rank_traveller()) %>% 
  dplyr::arrange(n_rank) %>% 
  dplyr::mutate(Winners = tidygraph::centrality_degree(mode = 'in'),
                bird_grp = dplyr::case_when(
                  grepl(pattern = "Penguin|penguin", x = name) ~ "Penguins",
                  grepl(pattern = "Kiwi", x = name) ~ "Kiwis",
                  grepl(pattern = "Albatross|Petrel|Shearwater|Gannet|Skua|Shag|Gull|Tern|Mollymawk", x = name) ~ "Seabirds",
                  grepl(pattern = "Dotterel|Stilt|Plover|Godwit|Oystercatcher|Wrybill", x = name) ~ "Shorebirds",
                  grepl(pattern = "Kākā|Kākāpō|Kea|Parakeet", x = name) ~ "Parrots",
                  grepl(pattern = "Harrier|Falcon|Owl|Morepork", x = name) ~ "Birds of Prey",
                  grepl(pattern = "Crake|Spoonbill|Heron|Bittern|Takahē|Weka|Pūkeko", x = name) ~ "Waders",
                  grepl(pattern = "Robin|Silvereye|Wren|Fantail|Warbler|Saddelback|Kōkako|Tūī|Bellbird|Tomtit|Mōhua|Rifleman|Stitchbird|Swallow|Fernbird|", x = name) ~ "Passerines",
                  grepl(pattern = "Teal|Duck|Dabchick", x = name) ~ "Water birds",
                  grepl(pattern = "Kererū", x = name) ~ "Pigeons",
                  grepl(pattern = "Cuckoo", x = name) ~ "Cuckoos"))


ggraph(graph, layout = 'fr') + 
    geom_edge_link(aes(colour = withingrp)) +
    geom_node_point(aes(size = Winners, colour = bird_grp)) + 
    scale_size_area(max_size = 10) +
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white') +
    labs(title = "A network of bird voters")

```

```{r}

ggraph(graph, "focus", focus = node_is_center()) +
  ggforce::geom_circle(aes(x0 = 0, yo = 0, r = r, data.frame(r = 1:5), colour = "grey")) +
  geom_edge_link() +
  geom_node_point() +
  coord_fixed()

```
