---
title: "Week46_12_11_19"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Week46_12_11_19}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}

library(ggExtra)
library(showtext)
library(ggplot2)
library(dplyr)
library(extrafont)

devtools::load_all()

```

Create defaults

```{r}

options(scipen = 100)

#Set ggplot theme
current_theme <- theme_classic() +
  theme(legend.position = c(0.65, 0.35),
        legend.direction = "horizontal",
        legend.margin = margin(0, 0, 0, -5),
        legend.justification = c(0, 0),
        legend.text = element_text(family = "Courier New", size = 12),
        legend.title = element_text(family = "Courier New", size = 15, face = "bold"),
        legend.text.align = 0,
        legend.spacing.x = unit(x = 0.25, units = "cm"),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(colour = "black", family = "Ubuntu", size = 15),
        axis.text.y = element_text(hjust = 2),
        axis.text.x = element_text(vjust = 4),
        axis.title = element_text(colour = "black", family = "Ubuntu", size = 20),
        plot.title = element_text(family = "Courier New", face = "bold", colour = "black",
                                  size = 20, hjust = 0.5),
        plot.subtitle = element_text(family = "Courier New", colour = "black",
                                     size = 15, hjust = 0.5),
        plot.caption = element_text(family = "Courier New", colour = "black",
                                     size = 12),
        plot.background = element_rect(fill = "white", size = 5),
        panel.background = element_rect(fill = "white"),
        plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "mm"))
  
  theme_set(current_theme)

```

```{r}

cran_code <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-12/loc_cran_packages.csv")

data("cran_dld_list_complete")

```

```{r, eval = FALSE}

unique_pkgs <- unique(cran_code$pkg_name)

library("dlstats")

pb <- dplyr::progress_estimated(n = length(unique_pkgs))

cran_dld_list_complete <- purrr::map(.x = unique_pkgs, .f = ~{
  
  pb$print()$tick()
  
  output <- tryCatch(cran_stats(..1), error = function(e){
    
    return(tibble::tibble(start = NA_character_,
                          end = NA_character_,
                          downloads = NA_integer_,
                          package = ..1))
  })
  
  return(output)
  
}) %>% 
  dplyr::bind_rows()

usethis::use_data(cran_dld_list, overwrite = TRUE)

```

# Determine total lines of code for all packages

Code impact method 1

In this method we just look at total downloads/line.

This will obviously bias longer-lived packages, but gives an idea of the total impact a package has had.

We need to know:

a) number lines of code
b) total downloads

```{r}

total_lines <- cran_code %>% 
  dplyr::group_by(pkg_name) %>% 
  dplyr::summarise(total_lines = sum(code))  %>% 
  #There are some cases where the number of lines of code are small
  #We exclude everything <2. 1 line seem to be datasets.
  dplyr::filter(total_lines > 1)

total_dlds <- cran_dld_list_complete %>% 
  dplyr::group_by(package) %>% 
  dplyr::summarise(downloads = sum(downloads))

code_impact <- dplyr::left_join(total_lines, total_dlds, by = c("pkg_name" = "package")) %>% 
  dplyr::mutate(code_impact = downloads/total_lines) %>% 
  dplyr::filter(total_lines < 50000)

```

Create arrow data for some of the best packages

```{r}

#Find highest impact package
hi_impact <- code_impact %>%
  dplyr::filter(code_impact == max(.$code_impact)) %>% 
  dplyr::mutate(text = "With 74 lines but >8 million downloads, the `dichromat` package has had the highest impact per line")

#Find package with most downloads
hi_dlds   <- code_impact %>%
  dplyr::filter(downloads == max(.$downloads)) %>% 
  dplyr::mutate(text = "Pipe is king! `magrittr` has had over 30 million downloads!")

```

Plot the code impact (i.e. download_rate ~ total_lines)

```{r}

ggplot(code_impact) +
  geom_point(aes(x = total_lines, y = downloads, fill = code_impact, size = code_impact), shape = 21, alpha = 0.5) +
  #Add magrittr curve/text
  geom_curve(data = hi_dlds,
             aes(xend = total_lines, yend = downloads, x = total_lines + 10000, y = downloads - 2000000),
             arrow = arrow(length = unit(0.1, units = "inches")), curvature = 0.1) +
  geom_text(data = hi_dlds,
            aes(x = total_lines + 15000, y = downloads - 3500000, label = stringr::str_wrap(text, 20)),
            size = 5, family = "Courier New") +
  #Add dichromat curve/text
  geom_curve(data = hi_impact,
             aes(xend = total_lines, yend = downloads, x = total_lines + 10000, y = downloads + 3000000),
             arrow = arrow(length = unit(0.1, units = "inches")), curvature = 0.1) +
  geom_text(data = hi_impact,
            aes(x = total_lines + 20200, y = downloads + 3700000, label = stringr::str_wrap(text, 30)),
            size = 5, family = "Courier New") +
  #Prevent clipping of the colourbar legend
  coord_cartesian(clip = "off") +
  #Add titles/caption
  labs(title = "Which R package has had the most impact?",
       subtitle = "Relationship between lines of code and total downloads",
       caption = "Plot: @ldbailey255 | Data: CRAN") +
  #Create blue colour gradient
  scale_fill_gradientn(colours = c("#004c6d", "#1b5a7d", "#2e698c", "#3e789d",
                                   "#4e87ad", "#5e96be", "#6ea6cf", "#7eb7e0", "#8ec7f2"),
                       na.value = "black", name = "Code impact \n (downloads/line)",
                       limits = c(0, 110000), breaks = seq(0, 100000, 25000),
                       labels = glue::glue('{number}K', number = seq(0, 100, 25))) +
  #Set x/y axes to show K/M abbreviations
  scale_x_continuous(breaks = seq(0, 50000, 10000),
                     labels = glue::glue('{number}K', number = seq(0, 50, 10)),
                     name = "Total lines (all languages)") +
  scale_y_continuous(breaks = seq(0, 30000000, 10000000),
                     labels = glue::glue('{number}M', number = seq(0, 30, 10)),
                     name = "Total downloads") +
  #Specify the guides
  guides(fill = guide_colorbar(direction = "horizontal",
                               barheight = unit(5, units = "mm"), 
                               barwidth = unit(100, units = "mm"),
                               draw.ulim = FALSE, 
                               ticks.colour = "transparent",
                               title.position = 'top',
                               title.hjust = 0.5,
                               title.vjust = 0.5,
                               label.hjust = 0.5,
                               label.vjust = 0.5),
         size = "none")

ggplot2::ggsave(filename = "../plots/Week46_2019.png", height = 10, width = 12, dpi = 300, type = "cairo")

```
