---
title: "Week45_2019_Nov5th"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Week45_2019_Nov5th}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We're going to create a map to show the most 'cycle crazed' cities in the US (measured as the % of individuals that cycle to work).

This will involve creating a map of the US with points for each city where the size is the total population and the fill is the % cyclists.

```{r}

#Load relevant packages
library(ggplot2)
#Used for scraping google maps info
library(ggmap)
#Using custom fonts
library(showtext)
library(sysfonts)
#palettes for plotting
library(rcartocolor)
#For spatial manipulation
library(sf)

```

```{r}

#Read in data
commute_mode <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-05/commute.csv")

#Read in custom font for plot
sysfonts::font_add_google("Supermercado One", "Supermercado One")
showtext_auto()
  
#Read in Google API
#Get one here https://developers.google.com/maps/documentation/maps-static/get-api-key
if(!ggmap::has_google_key()){
  
  ggmap::register_google(key = readline("Enter your Google API:"))
  
}
  
#Use google API to scrape location data of each city
#Takes a bit of time!!
plot_data <- commute_mode %>%
  #Filter out bike data
  dplyr::filter(mode == "Bike" & !state_abb %in% c("AK", "HI")) %>% 
  #Combine city and state data together for web scraping
  dplyr::mutate(city_state = paste(city, state, sep = ", ")) %>% 
  #This will add lat/long for each city from google maps
  ggmap::mutate_geocode(city_state)

```

Create a square grid of cycling cities across the US

```{r}

buffer_size <- tibble::tibble(city_size = c("Large", "Medium", "Small"),
                              buffer = c(30000, 15000, 10000))

#Firstly, turn all our data into circle polygons
plot_data_sf <- sf::st_as_sf(plot_data, coords = c("lon", "lat"), crs = CRS("+init=epsg:4326")) %>% 
  #Transform to planar coordinate system
  sf::st_transform(5070) %>% 
  #Split into small, medium, large city size
  split(., f = as.factor(.$city_size)) %>% 
  purrr::map2(.x = ., .y = names(.), .f = function(data, city_size, buffer_size){
    
    return(sf::st_buffer(data, dist = buffer_size[which(buffer_size$city_size == city_size), ]$buffer))
    
  }, buffer_size)

#Combine back together
plot_data_sf <- do.call(rbind, plot_data_sf) %>% 
  sf::st_cast("MULTIPOLYGON")

ggplot()+
    geom_sf(data = plot_data_sf)

```

Read in US map

```{r}

#Read in data and make it an sf object
us_map <- sf::st_as_sf(maps::map('usa', plot = FALSE, fill = TRUE), crs = CRS("+init=epsg:4326")) %>% 
  #Make into same CRS
  sf::st_transform(5070) %>% 
  #Simplify and buffer slightly to prevent errors
  sf::st_simplify(dTolerance = 0.0001) %>% 
  sf::st_buffer(dist = 0)

us_states <- sf::st_as_sf(maps::map('state', plot = FALSE, fill = TRUE), crs = CRS("+init=epsg:4326")) %>% 
  #Make into same CRS
  sf::st_transform(5070) %>% 
  #Simplify and buffer slightly to prevent errors
  sf::st_simplify(dTolerance = 0.0001) %>% 
  sf::st_buffer(dist = 0)

ggplot()+
  geom_sf(data = us_map) +
  geom_sf(data = plot_data_sf)

```

Make square grid of polygon

```{r}

us_map_grid <- us_map %>% 
  sf::st_make_grid(cellsize = 25000) %>%
  sf::st_intersection(us_map) %>% 
  sf::st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(id = row_number())

ggplot()+
  geom_sf(data = us_map_grid) +
  geom_sf(data = us_map, fill = NA)

```

Spatial join polygons

```{r}

#Determine mean cycle percentage in each grid cell
us_point_join <- sf::st_join(us_map_grid, plot_data_sf) %>% 
  dplyr::group_by(id) %>% 
  dplyr::summarise(mean_cycle = mean(percent, na.rm = TRUE))

ggplot()+
  geom_sf(data = us_point_join, aes(fill = mean_cycle)) +
  labs(title = paste0(emo::ji("bicycle"), " Cycling in the US ", emo::ji("bicycle")),
       subtitle = "Where do people cycle to work most?") +
  scale_fill_gradientn(colours = c("#217a79",
"#348c82",
"#499d89",
"#60af8f",
"#7ac094",
"#95d199",
"#b3e29e",
"#d3f2a3"),
                       na.value = "black", name = "Avg. percentage of cyclists",
breaks = seq(0, 12, 2), limits = c(0, 12)) +
  guides(fill = guide_colorbar(direction = "horizontal",
                               barheight = unit(10, units = "mm"), 
                               barwidth = unit(100, units = "mm"),
                               draw.ulim = FALSE, 
                               ticks.colour = "transparent",
                               title.position = 'top',
                               title.hjust = 0.5,
                               title.vjust = -6,
                               label.hjust = 0.5,
                               label.vjust = 7)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(family = "Supermercado One", size = 100, hjust = 0.5, vjust = 0),
        plot.subtitle = element_text(family = "Supermercado One", size = 80, hjust = 0.5, vjust = 0),
        plot.margin = margin(t = 0, r = 2, b = 2, l = 2, unit = "cm"),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        legend.background = element_blank(),
        legend.position = c(0.2, 0.11),
        legend.title = element_text(family = "Supermercado One", size = 70),
        legend.text = element_text(family = "Supermercado One", size = 50),
        legend.margin = margin(t = 0, b = 0, r = 3, l = 3, unit = "mm"))

ggplot2::ggsave(filename = "./plots/5_11_2019.png", height = 10, width = 15)

```

Other plot ideas that I didn't like

```{r}

plot_data$city_size <- forcats::fct_relevel(plot_data$city_size, c("Small", "Medium", "Large"))

plot_data_city <- plot_data %>% 
  dplyr::filter(city_size %in% c("Large", "Medium"))

best_cities <- plot_data_city %>% 
  arrange(-percent) %>% 
  slice(1:3)

us_map <- ggplot2::map_data("state")

ggplot()+
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group),
               fill = "white", colour = "black") +
  geom_point(data = plot_data_city, aes(x = lon, y = lat,
                                   size = city_size, fill = percent),
             shape = 21) +
    scale_fill_carto_c(type = "quantitative", palette = "BluGrn", direction = -1,
                       name = "Percentage \n of cyclists") +
  scale_size_discrete(guide = FALSE, range = c(4, 6)) +
  guides(fill = guide_colorbar(direction = "horizontal",
                               barheight = unit(7, units = "mm"), 
                               barwidth = unit(100, units = "mm"),
                               draw.ulim = FALSE, 
                               ticks.colour = "transparent",
                               title.position = 'top',
                               title.hjust = 0.5,
                               title.vjust = 0.25,
                               label.hjust = 0.5)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 2, l = 2, unit = "cm"),
        plot.background = element_rect(fill = "#D4D2D5"),
        panel.background = element_rect(fill = "#D4D2D5"),
        legend.background = element_rect(fill = "#D4D2D5", colour = "black"),
        legend.position = c(0.15, 0.1),
        legend.title = element_text(family = "Supermercado One", size = 25),
        legend.text = element_text(family = "Supermercado One", size = 15),
        legend.margin = margin(t = 3, b = 3, r = 3, l = 3, unit = "mm"))

ggplot2::ggsave(filename = "../plots/5_11_2019.pdf", height = 8, width = 12)

```

State level plot

```{r}

plot_data_state <- plot_data %>% 
  dplyr::group_by(state) %>% 
  dplyr::summarise(mean_perc = mean(percent),
                   state_abb = first(state_abb)) %>% 
  dplyr::mutate(state = dplyr::recode(toupper(state), CA = "CALIFORNIA"))

#Add to polygon
us_map <- ggplot2::map_data("state") %>% 
  dplyr::mutate(region = toupper(region)) %>% 
  dplyr::left_join(plot_data_state, by = c("region" = "state"))

#Determine the midpoint of each state to add state abbreviations
state_centres <- split(us_map, as.factor(us_map$region)) %>% 
  purrr::map(.f = ~{Polygon(coords = ..1[, c("long", "lat")])}) %>% 
  SpatialPolygons(Srl = Polygons(.), proj4string = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
  
ggplot()+
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group,
                                  fill = mean_perc), colour = "black") +
  scale_fill_carto_c(type = "quantitative", palette = "BluGrn", direction = -1,
                       name = "Percentage \n of cyclists") +
  guides(fill = guide_colorbar(direction = "horizontal",
                               barheight = unit(7, units = "mm"), 
                               barwidth = unit(100, units = "mm"),
                               draw.ulim = FALSE, 
                               ticks.colour = "transparent",
                               title.position = 'top',
                               title.hjust = 0.5,
                               title.vjust = 0.25,
                               label.hjust = 0.5)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 2, l = 2, unit = "cm"),
        plot.background = element_rect(fill = "#D4D2D5"),
        panel.background = element_rect(fill = "#D4D2D5"),
        legend.background = element_rect(fill = "#D4D2D5", colour = "black"),
        legend.position = c(0.15, 0.1),
        legend.title = element_text(family = "Supermercado One", size = 25),
        legend.text = element_text(family = "Supermercado One", size = 15),
        legend.margin = margin(t = 3, b = 3, r = 3, l = 3, unit = "mm"))

```